---
title: "Video game sales"
output: html_document
date: "2023-03-18"



## 1. Scenerio
            
Suppose their is an investor who wants to invest in a gaming platform. However before the       investor invests, they want to know which gaming platform between the years 2010 to 2016 has the best NA, EU, and Global sales. The investor also wants to know what video game category produces the most income between 2010 and 2016 while for each video game platform so that they can better invest their money to better profit themselves in the future. The investor also wants to see what developers create more profitable games and the video game categories those developers mainly develop. The investor wants an in-depth analyst on the video game platforms, video game categories, and developers before they invest their money. 

## 2.Preparation
  The data we are using was uploaded to Kaggle by SID_TWR. SID_TWR stated that this dataset was      scraped from VGChartz and Metacritic.

2.1 Load Tidyverse package
Before we work on our data we first want to load it into R, but first we must add two R packages into R before we attempt to work with data.

```{r} #install.packages("Tidyverse")
  library(tidyverse) #Used to load tidyverse package into R.
```


2.2 Import dataset  
Now we want to use import our dataset into R using the read_CSV function that is from our readr package.
```{r}
Video_Games_Sales <- read_csv("C:/Users/Matthew Afoakwah/OneDrive/Documents/Bport/Bport Spring 2023 (Final Semester!)/CSC435 Data Analytics/Project/CSC435_Video_Games_Sales_Project/data/Video_Games_Sales_as_at_22_Dec_2016.csv") #Reads the CSV file from the file location and imports the file into R, named Video_Games_Sales
```

2.3 Preview dataset
Before we can work on our dataset, we must first see if the dataset is relivant to our task. In order to accomplish this we must view the dataset in R.
```{r}
View(Video_Games_Sales) #view() function shows the entire dataset.
```

```{r}
head(Video_Games_Sales)
```



```{r}
str(Video_Games_Sales)
```

```{r}
colnames(Video_Games_Sales)
```
```{r}
length(Video_Games_Sales)
```

```{r}
dim(Video_Games_Sales)
```

## 3. Data Cleaning

We will be using cleaning data using data transformations.

3.1 Since we are focusing on the video game sales between 2010 to 2016 we will shrink the data set by filtering it.

```{r}
gamesales2010_2016 <- filter(Video_Games_Sales, Year_of_Release >= 2010)
nrow(gamesales2010_2016)
```

3.2 We will now filter all rows where the Year_of_Release is unknown.

```{r}
gamesales2010_2016 <- filter(gamesales2010_2016, Year_of_Release != "N/A")
nrow(gamesales2010_2016)
```

## 4. Data Exploration
We now develop questions to explore the data set.
  * Which video game platform made the most NA Sales from 2010 to 2016?
  * Which video game platform made the most Global Sales from 2010 to 2016?
  * Which video game category had the highest NA sales?
  * Which video game category made the highest Global sales?
  * Which top 5 video game developers have the highest Global video game sales?
  * Which video game genre is the top three developers most profitable in

4.1 To answer the first question we have to calculate the NA revenue of each platform.

```{r}
platform_NA_Sales <- gamesales2010_2016 %>%
  group_by(Platform) %>%
  summarize(NA_total = sum(NA_Sales)) %>%
  arrange(NA_total)
  view(platform_NA_Sales)
```

4.2 Secondly, we calculate the global revenue for each platform.
```{r}
platform_glbl_sales <- gamesales2010_2016 %>%
  group_by(Platform) %>%
  summarize(glbl_total = sum(Global_Sales)) %>%
  arrange(glbl_total)
view(platform_glbl_sales)
```


4.3 Then, we calculate the NA revenue based on categories.
```{r}
cat_na_sales <- gamesales2010_2016 %>%
  group_by(Genre) %>%
  summarise(cat_na_total = sum(NA_Sales)) %>%
  arrange(cat_na_total)
view(cat_na_sales)
```


4.4 Now, we calculate the Global revenue based on categories.
```{r}
cat_glbl_sales <-gamesales2010_2016 %>%
  group_by(Genre) %>%
  summarise(cat_glbl_total = sum(Global_Sales)) %>%
  arrange(cat_glbl_total)
view(cat_glbl_sales)
```


4.5 After, we want to calculate the global revenue based on the developer.
```{r}
dev_global <- gamesales2010_2016 %>%
  filter(Developer != "N/A") %>%
  group_by(Developer) %>%
  summarise(global_total = sum(Global_Sales)) %>%
  arrange(-global_total) %>%
  slice_head(n = 5)
view(dev_global)
```



4.6 Finally, we want to find what video game categories are mainly developed by the top three developers and how the revenue in these categories differ.
```{r}
dev_genre_global <- gamesales2010_2016 %>%
  filter(Developer != "N/A") %>%
  filter(Genre != "N/A") %>%
  group_by(Genre, Developer) %>%
  summarise(devcat_total = sum(Global_Sales)) %>%
  arrange(-devcat_total)
view(dev_genre_global)
```





